Predicting the Ending Notes of Bach Chorales

### 2018-04-25 Paper.tex
Added note about KNN model

\subsection{Computer Model}
In the final model, each chorale was used in a decision tree to predict the final pitch. 80 random chorales from the data set were used as the training set, with the remaining 20 chorales left for testing. Five total trees were created, with a different subset of 20 each time; these five trees therby predicted over the entire data set. This model was able to correctly predict 55 of the 100 chorales. Looking at the various decision trees generated by the model, most trees contained the same node: checking whether the starting key was either A or E major. In some cases, there would be only one solution for chorales which were in A and E major, usually MIDI note 69 (A4). Overall, the most prevalent node checked the starking key signature. After key signature, the most prevalent nodes checked the pitch of the first note. While these covered the majority of the nodes within each tree, there were a number of outlier nodes. For instance, one fairly common outlier was built on the pitch of the 16th or 17th note in the melody line; this note is likely the end of the first phrase of the melody. Other nodes were related to the overall speed of the chorale; one node checked the length of the 15th note, and another checked if the 34th note of the melody was on the 34th beat or later. There may be a pattern apparent in the data, where faster chorales are more likely to end on a particular note than slower chorales.\\
\begin{figure}[h]
  \includegraphics[width=13.5cm]{Tree5}
  \centering
  \caption{One of the five decision trees generated by the computer model, with the nodes relabeled for readability. This tree was trained on chorales 1 through 80, and tested on chorales 81 through 100.}
\end{figure}

Although the human and computer models performed at similar levels, one noticeable difference between the two is accuracy. The confusion matrix for the computer model had fewer rows and columns than the human model. The pitches for the notes were given in a range between 60 and 75 in the MIDI scale, but this range applies to the entire melody line, not the endings. While the human model did not assume any distribution of pitches in the range, the computer model determined that the ending notes are clustered in the middle of the range, thus explaining the shrinkage of the confusion matrix.
\section{Conclusion}
Although the computer model performed at a similar level to the human model
Regardless of the results of the model, the relationship between computers and music continues to provide a challenging roadblock on the path to human-like artificial intelligence. The amount of interdisciplinary cooperation that will eventually be necessary is staggering; from computer scientists and mathematicians to create the models, to musicians for the actual art, to psychologists and philosophers to determine how a computer can best ``learn.''

\begin{thebibliography}{10}
\bibitem{computer_composer}
  De Felice, Clelia, et al. "Splicing music composition." Information Sciences, vol. 385 - 386, 2017, pp. 196-212. OhioLINK Electronic Journal Center, doi:10.1016/J.INS.2017.01.004.

>>>>

\subsection{Computer Model}
In the final model, each chorale was used in a decision tree to predict the final pitch. 80 random chorales from the data set were used as the training set, with the remaining 20 chorales left for testing. Five total trees were created, with a different subset of 20 each time; these five trees therby predicted over the entire data set.\\

This model was able to correctly predict 55 of the 100 chorales. 21 chorales were overestimated, and 24 chorales were underestimated. Looking at the various decision trees generated by the model, most trees contained the same node: checking whether the starting key was either A or E major. In some cases, there would be only one solution for chorales which were in A and E major, usually MIDI note 69 (A4). Overall, the most prevalent node checked the starking key signature. After key signature, the most prevalent nodes checked the pitch of the first note. While these covered the majority of the nodes within each tree, there were a number of outlier nodes. For instance, one fairly common outlier was built on the pitch of the 16th or 17th note in the melody line; this note is likely the end of the first phrase of the melody. Other nodes were related to the overall speed of the chorale; one node checked the length of the 15th note, and another checked if the 34th note of the melody was on the 34th beat or later. There may be a pattern apparent in the data, where faster chorales are more likely to end on a particular note than slower chorales.\\
\begin{figure}[h]
  \includegraphics[width=13.5cm]{Tree5}
  \centering
  \caption{One of the five decision trees generated by the computer model, with the nodes relabeled for readability. This tree was trained on chorales 1 through 80, and tested on chorales 81 through 100.}
\end{figure}

Although the human and computer models performed at similar levels, one noticeable difference between the two is accuracy. The confusion matrix for the computer model had fewer rows and columns than the human model. The pitches for the notes were given in a range between 60 and 75 in the MIDI scale, but this range applies to the entire melody line, not the endings. While the human model did not assume any distribution of pitches in the range, the computer model determined that the ending notes are clustered in the middle of the range, thus explaining the shrinkage of the confusion matrix.

A similar process was used with a $k$ nearest neighbors approach, but the results were significantly worse than the classification tree. With $k=1$, only 13 chorales were predicted correctly; increasing the value of $k$ was only able to raise the correct predictions to the mid-20s.
\section{Conclusion}
Although the computer model performed at a similar level to the human model
Regardless of the results of the model, the relationship between computers and music continues to provide a challenging roadblock on the path to human-like artificial intelligence. The amount of interdisciplinary cooperation that will eventually be necessary is staggering; from computer scientists and mathematicians to create the models, to musicians for the actual art, to psychologists and philosophers to determine how a computer can best ``learn.'' 
\newpage
\begin{thebibliography}{10}
\bibitem{computer_composer}
  De Felice, Clelia, et al. "Splicing music composition." Information Sciences, vol. 385 - 386, 2017, pp. 196-212. OhioLINK Electronic Journal Center, doi:10.1016/J.INS.2017.01.004.
